Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-19 numerical-starformation.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

Modeling star formation in the context of galaxy formation forces one
to confront two fundamental problems. The dynamical range required to
study galaxy formation and evolution is substantial, and attempting to
directly resolve the formation of individual stars in the context of a
Milky Way-type galaxy is computationally unfeasible.  Also, a strong
theoretical understanding of the details of star formation is
currently lacking.  Due to these challenges, models of star formation
in cosmological simulations are largely phenomenological, and the
``stars'' that are formed are in most cases meant to represent an
ensemble stellar population (although not all; see
Section~\ref{sec:starform_pop3}).

The \enzo\ code includes approximately ten models for star formation
and feedback.  Broadly speaking, these methods all work in similar
ways: At a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., that have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but may include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with
the same mass is placed in the center of that cell, with the same
momentum vector as the removed gas.  This star particle is then
allowed to inject mass, momentum, thermal energy, metals, and possibly
magnetic fields and/or cosmic ray populations into its local
environment.  In general, the particle is treated as an ensemble of
stars, with feedback properties occuring over time according to the
assumed initial mass function of the stellar population.

@@

Algorithmically, the star formation events in Kravtsov's code are
assumed to occur once every global time step (with the constraint $\Delta t_0 \leq
10^7$ years).
In cells where star formation is determined to occur (i.e. $\rho_{\rm
  gas} \geq \rho_{\rm SF}$), star particles with a mass of $m_* =
\dot{\rho}_* V_{\rm cell} \Delta t_0$ (where $V_{\rm cell}$ is the
volume of the mesh cell) are assumed to form instantaneously in a
manner similar to that described in Section~\ref{sec:starform_cen}.
The \enzo\ implementation of this algorithm is similar, except that
instead of forming stars only at the root grid time step, we allow
stars to form at the time step of the highest level of resolution at
any particular point in space.  As can be seen from the equation for
$m_*$ above, this can result in very small stellar masses.  To avoid
memory and processor time issues related to having very large numbers
of star particles we impose a threshold mass $M_{\rm *,min}$ such that
a star particle only forms if $m_* \geq M_{\rm *,min}$.  An
appropriate choice of $M_{\rm *,min}$ does not significantly change
the star overall star formation history of a simulation, though it may
delay the onset of star formation in a given cell relative to a
simulation without a particle mass threshold.

@@

Kravtsov reports that simulations with this algorithm reliably
reproduce the star formation rate-gas surface density relation of the
Schmidt law, and are not particularly sensitive to numerical
parameters \citep{2003ApJ...590L...1K}.  He also notes that this
algorithm is surprisingly insensitive to the presence or absence of
feedback and details of the cooling and heating properties of the gas,
which suggests that the global star formation rate is determined by
gravitationally- driven supersonic turbulence (on large scales) rather
than stellar feedback or thermal instabilities on small scales.

>>>>

% Modeling star formation in the context of galaxy formation forces one
% to confront two fundamental problems. The dynamical range required to
% study galaxy formation and evolution is substantial, and attempting to
% directly resolve the formation of individual stars in the context of a
% Milky Way-type galaxy is computationally unfeasible.  Also, a strong
% theoretical understanding of the details of star formation is
% currently lacking.  Due to these challenges, models of star formation
% in cosmological simulations are largely phenomenological, and the
% ``stars'' that are formed are in most cases meant to represent an
% ensemble stellar population (although not all; see
% Section~\ref{sec:starform_pop3}).

Due to the computationally unfeasible number of stars in a galaxy
(10^11) and the lack of detailed understanding of star formation, a
number of phenomenological star formation models are included in
\enzo.
The \enzo\ code includes approximately ten models for star formation
and feedback.  Broadly speaking, these methods all work in similar
ways: At a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., that have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but may include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with
the same mass is placed in the center of that cell, with the same
momentum vector as the removed gas.  This star particle is then
allowed to inject mass, momentum, thermal energy, metals, and possibly
magnetic fields and/or cosmic ray populations into its local
environment.  In general, the particle is treated as an ensemble of
stars, with feedback properties occuring over time according to the
assumed initial mass function of the stellar population.

@@

Algorithmically, the star formation events in the Scmidt-law algorithm are
assumed to occur once every global time step (with the constraint $\Delta t_0 \leq
10^7$ years).
In cells where star formation is determined to occur (i.e. $\rho_{\rm
  gas} \geq \rho_{\rm SF}$), star particles with a mass of $m_* =
\dot{\rho}_* V_{\rm cell} \Delta t_0$ (where $V_{\rm cell}$ is the
volume of the mesh cell) are assumed to form instantaneously in a
manner similar to that described in Section~\ref{sec:starform_cen}.
The \enzo\ implementation of this algorithm is similar, except that
instead of forming stars only at the root grid time step, we allow
stars to form at the time step of the highest level of resolution at
any particular point in space.  As can be seen from the equation for
$m_*$ above, this can result in very small stellar masses.  To avoid
memory and processor time issues related to having very large numbers
of star particles we impose a threshold mass $M_{\rm *,min}$ such that
a star particle only forms if $m_* \geq M_{\rm *,min}$.  An
appropriate choice of $M_{\rm *,min}$ does not significantly change
the star overall star formation history of a simulation, though it may
delay the onset of star formation in a given cell relative to a
simulation without a particle mass threshold.

@@

% Kravtsov reports that simulations with this algorithm reliably
% reproduce the star formation rate-gas surface density relation of the
% Schmidt law, and are not particularly sensitive to numerical
% parameters \citep{2003ApJ...590L...1K}.  He also notes that this
% algorithm is surprisingly insensitive to the presence or absence of
% feedback and details of the cooling and heating properties of the gas,
% which suggests that the global star formation rate is determined by
% gravitationally- driven supersonic turbulence (on large scales) rather
% than stellar feedback or thermal instabilities on small scales.

### 2013-05-19 numerical-timestep.tex
Rearranging sections.
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

\subsection{Time Stepping}
\label{sec.timestepping}

@@

In Equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $\min ( \ldots
)_L$ formalism means that this value is calculated for all cells or
particles on a
given level L and the minimum overall value is taken as the timestep.
Equation~\ref{eqn:dthydro} ensures that all cells satisfy the
Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of
an explicit finite difference discretization of the Euler equations.
Effectively this condition forces the timestep to be small enough such
that any changes in the fluid propagate less than a single grid
spacing, $\Delta x$.  In this equation, $\kappa_{\rm hydro}$ is a
numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 1$
(with a typical value of $\kappa_{\rm hydro} \sim 0.3-0.5$) that
ensures that the CFL condition is always met, and $c_s$ and $v_x$ are
the sound speed and peculiar baryon velocity in a given cell.

>>>>

\section{Time Stepping}
\label{sec.timestepping}

@@

In Equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $\min ( \ldots
)_L$ formalism means that this value is calculated for all cells or
particles on a
given level L and the minimum overall value is taken as the timestep.
Equation~\ref{eqn:dthydro} ensures that all cells satisfy the
Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of
an explicit finite difference discretization of the Euler equations.
Effectively this condition forces the timestep to be small enough such
that any changes in the fluid propagate less than a single grid
spacing, $\Delta x$.  In this equation, $\kappa_{\rm hydro}$ is a
numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 0.5$
(with a typical value of $\kappa_{\rm hydro} \sim 0.3-0.5$) that
ensures that the CFL condition is always met, and $c_s$ and $v_x$ are
the sound speed and peculiar baryon velocity in a given cell.

### 2013-05-19 overview.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

The total fluid energy $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a}
\end{equation}
The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
%
\begin{eqnarray}
%
e   & = & p / [(\gamma - 1) \rho ],
        \label{eq:eq_of_state} \\
%
\nabla^2 \phi & = & 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
        \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
Equation (\ref{eq:potential}) includes contributions to the
gravitational potential $\phi$ from the total mass density $\rho_{\rm
  total}$.  Although we write the equations including both MHD and
comoving coordinates, the code is frequently used both in the pure
hydrodynamic limit ($\vecB = 0$) and without cosmological expansion
($a = 1$, $\dot{a} = 0$).

@@

The radiative transfer equation in comoving coordinates 
\citep[e.g.,][]{Gnedin97} reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
specific intensity in units of energy per time $t$ per solid angle per
unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
ratio of scale factors at the current time and time of emission.  The
second term represents the propagation of radiation, where the factor
$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
both the cosmological redshift and dilution of radiation.  On the
right hand side, the first term considers the absorption coefficient
$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
that includes any point sources of radiation or diffuse radiation.

@@

In the other limit, we solve an equation derived from integrating
the equation of radiative transfer over angles.  In particular, we use a
flux-limited diffusion approximation for radiative
transfer, with couplings to both the gas energy and chemical number
densities. The equations solved are:
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
    \nabla\cdot\left(D\nabla E_r\right) -
    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r$ is a grey radiation energy density and $e_c$ is the
internal energy correction due to photo-heating and chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define the grey radiation energy density
through first assuming a fixed frequency spectrum, i.e.
$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
defining the integrated quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
square-root flux-limiter} \citep[see][]{Morel2000}, $\kappa$ is the total
opacity, $\eta$ is the field of radiation sources, $\Gamma$ provides the
radiation induced photo-heating, and $\Lambda$ the chemical
cooling.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

The grid hierarchy begins with the root grid, which covers the entire
domain of interest with a coarse, uniform, Cartesian grid. Then, as
the solution evolves and interesting regions form, finer meshes are
placed below these regions (we use the notation `below' to refer to
finer grids and `above' for coarser grids).  We restrict the ratio
between cell sizes to be an integer, typically 2 or 4, and refer to a
level as all grids with the same cell size.  In order to keep
things as simple as possible, the edges of subgrids must coincide with
the cell edge of its immediate parent (coarser) grid. Additionally,
the hierarchy can be initialized with one or more static grids if a
higher initial resolution is required.

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum time step allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1$
($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the allowed
timesteps for those grids.  If there are more levels, we repeat this
procedure until the bottom level of the hierarchy has been reached.
Once there, we continue advancing the grids on the lowest level until
they have `caught up' to the next highest level above (i.e. $\sum
\Delta t_l = \Delta t_{l-1}$).  This procedure repeats itself until
all grids have been advanced by a total time of $\Delta t_0$.

Since interesting regions on the grid may move, the hierarchy must
adapt itself.  We do this whenever a level has caught up to the
coarser level above it and consists of entirely rebuilding the grids
on that level and below.  This is done by applying the grid refinement
criteria to the grids on that level, flagging zones that require
extra grids.  These criteria depend on the physical problem being
simulated.  We have implemented a number of options, including shock
and steep gradient detectors, but in the problem described here,
employ one based on the mass within a cell, imitating the Lagrangian
nature of the SPH algorithm.  Once a grid has a set of flagged cells,
we run a machine-vision based algorithm \citep{Berger91} to find edges
and determine a good placement of subgrids.  These subgrids must not
overlap one another, must cover all flagged cells and their
neighbouring cells, and be above a preset efficiency threshold, where
the efficiency of a cell is defined as the ratio of flagged cells to
total cells.  Once these new subgrids have been identified, the
solution from the next coarser grid is interpolated (see below) in
order to initialize the values on the new grids.  Finally, any overlap
between these new subgrids and the old ones is identified and the
solution within the regions of overlap is copied to the new subgrids.
The entire procedure just outlined is then repeated on the new grids
and in this way the entire hierarchy (from the original level examined
and below) is rebuilt.

@@

The current implementation of self-gravity in \enzo\ uses a Fast
Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
top (level 0) grid.  The advantage of using this method is that it
naturally allows both periodic and isolated boundary conditions for
the gravity, choices which are very common in astrophysics and
cosmology.  On subgrids, we interpolate the boundary conditions from
that grid's parent (either the level 0 grid or some other subgrid).
The Poisson equation is then solved using a multigrid technique.  In
\enzo, self-gravity is optional.  Aside from calculating the
gravitational potential from the baryon fields and particles existing
in the simulation, there are also a number of options for static gravitational fields.

@@

A second option for radiative transfer is a moment-based method that
adds an additional field tracking the radiant energy density
This field is evolved using the flux-limited diffusion method, which
transitions smoothly between streaming (optically thin) and opaque
limits and is coupled to the hydrogen ionization network.  The
resulting set of linear equations is solved using the parallel HYPRE
framework.  Full details for the enzo implementation can be found in
\citet{ReynoldsHayesPaschosNorman2009}.

>>>>

The total fluid energy $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a}
\end{equation}
The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
%
\begin{eqnarray}
%
e   & = & p / [(\gamma - 1) \rho ],
        \label{eq:eq_of_state} \\
%
\nabla^2 \phi & = & 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
        \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
Equation (\ref{eq:potential}) includes contributions to the
gravitational potential $\phi$ from the total mass density $\rho_{\rm
  total}$, which is defined as $\rho_{\rm gas} + \rho_{\rm dm} + \rho_{\rm stars}
- \rho_{\rm average}$.  Although we write the equations including both MHD and
comoving coordinates, the code is frequently used both in the pure
hydrodynamic limit ($\vecB = 0$) and without cosmological expansion
($a = 1$, $\dot{a} = 0$).

@@

The radiative transfer equation in comoving coordinates 
\citep[e.g.,][]{Gnedin97} reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
specific intensity in units of energy per time $t$ per solid angle per
unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
ratio of scale factors at the current time and time of emission.  The
second term represents the propagation of radiation, where the factor
$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
both the cosmological redshift and dilution of radiation.  On the
right hand side, the first term considers the absorption coefficient
$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
that includes any sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, we solve an equation derived from integrating
the equation of radiative transfer over angles.  In particular, we use a
flux-limited diffusion approximation for radiative
transfer, with couplings to both the gas energy and chemical number
densities. The equations solved are:
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
    \nabla\cdot\left(D\nabla E_r\right) -
    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \nabla\cdot\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(x,t)$ is a grey radiation energy density and $e_c$ is the
internal energy correction due to photo-heating and chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define the grey radiation energy density
through first assuming a fixed frequency spectrum, i.e.
$E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t) \chi(\nu)$, and then
defining the integrated quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
square-root flux-limiter} \citep[see][]{Morel2000}, $\kappa$ is the total
opacity, $\eta$ is the field of radiation sources, $\Gamma$ provides the
radiation induced photo-heating, and $\Lambda$ the chemical
cooling.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

The grid hierarchy begins with the root grid, which covers the entire
domain of interest with a coarse, uniform, Cartesian grid. Then, as
the solution evolves and interesting regions form, finer meshes are
placed within these coarse regions.
% (we use the notation `below' to refer to
% finer grids and `above' for coarser grids). 
 We restrict the ratio
between cell sizes to be an integer, typically 2 or 4, and refer to a
level as all grids with the same cell size.  In order to keep
things as simple as possible, the edges of subgrids must coincide with
the cell edge of its immediate parent (coarser) grid. Additionally,
the hierarchy can be initialized with one or more static grids if a
higher initial resolution is required.

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum time step allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1$
($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the allowed
timesteps for those grids.  If there are more levels, we repeat this
procedure until the bottom level of the hierarchy has been reached.
Once there, we continue advancing the grids on the lowest level until
they have `caught up' to the coarser parent grids (i.e. $\sum
\Delta t_l = \Delta t_{l-1}$).  This procedure repeats itself until
all grids have been advanced by a total time of $\Delta t_0$ (see
Figure~\ref{fig:wcycle} for a schematic).

Since interesting regions on the grid may move, the hierarchy must
adapt itself.  We do this whenever a level has caught up to the
coarser parent level and consists of entirely rebuilding the grids
on that level and at finer resolutions.  This is done by applying the grid refinement
criteria to the grids on that level, flagging zones that require
extra grids.  These criteria depend on the physical problem being
simulated.  We have implemented a number of options, including shock
and steep gradient detectors, but in the problem described here,
employ one based on the mass within a cell, imitating the Lagrangian
nature of the SPH algorithm.  Once a grid has a set of flagged cells,
we run a machine-vision based algorithm \citep{Berger91} to find edges
and determine a good placement of subgrids. 
%  These subgrids must not overlap one another, must cover all flagged
% cells and their neighbouring cells, and be above a preset efficiency
% threshold, where the efficiency of a cell is defined as the ratio of
% flagged cells to total cells. 
Once these new subgrids have been identified, the
solution from the next coarser grid is interpolated (see below) in
order to initialize the values on the new grids.  Finally, any overlap
between these new subgrids and the old ones is identified and the
solution within the regions of overlap is copied to the new subgrids.
The entire procedure just outlined is then repeated on the new grids
and in this way the entire hierarchy (from the original level examined
and continuing on to finer levels) is rebuilt.

@@

The current implementation of self-gravity in \enzo\ uses a Fast
Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
coarsest (Level 0) grid.  The advantage of using this method is that
it is fast, accurate, and 
naturally allows both periodic and isolated boundary conditions for
the gravity, choices which are very common in astrophysics and
cosmology.  On subgrids, we interpolate the boundary conditions from
that grid's parent (either the level 0 grid or some other subgrid).
The Poisson equation is then solved using a multigrid technique.  In
\enzo, self-gravity is optional.  Aside from calculating the
gravitational potential from the baryon fields and particles existing
in the simulation, there are also a number of options for static gravitational fields.

@@

A second option for radiative transfer is a moment-based method that
adds an additional field tracking the radiant energy density.
This field is evolved using the flux-limited diffusion method, which
transitions smoothly between streaming (optically thin) and opaque
limits and is coupled to the hydrogen ionization network.  The
resulting set of linear equations is solved using the parallel HYPRE
framework.  Full details for the enzo implementation can be found in
\citet{ReynoldsHayesPaschosNorman2009}.
